---
title: "Statistical Inference, Data Analysis Part 2"
author: "Jackson Pham"
date: "`r Sys.Date()`"
output: html_document
---

```{r, echo = TRUE}

# Load necessary libraries
library(datasets)
library(psych)

```

```{r, echo = TRUE}

# Load ToothGrowth dataset
data <- datasets::ToothGrowth

# Summarize statistic
summary(data)

# Describe the data
describe(data)

```

```{r, echo = TRUE}

# Visualize the data
boxplot(len ~ dose * supp, data = ToothGrowth, xlab = "Dose", ylab = "Tooth Length", main = "Tooth Growth by Dose and Supplement")

# Calculate t.test between OJ and VC
t_test_result <- t.test(len ~ supp, data = data)
print(t_test_result)

# Subset the data for 'VC' group
vc_data <- subset(data, supp == "VC")
# Subset the data for 'OJ' group
oj_data1 <- subset(data, supp == "OJ")

# Subset the data further for VC dose levels 0.5 and 1.0
vc_dose_0.5 <- subset(vc_data, dose == 0.5)$len
vc_dose_1.0 <- subset(vc_data, dose == 1.0)$len
vc_dose_2.0 <- subset(vc_data, dose == 2.0)$len

# Subset the data further for dose levels 0.5 and 1.0
oj_dose_0.5 <- subset(oj_data1, dose == 0.5)$len
oj_dose_1.0 <- subset(oj_data1, dose == 1.0)$len
oj_dose_2.0 <- subset(oj_data1, dose == 2.0)$len

# Perform the t-test VC
t_test_result1 <- t.test(vc_dose_0.5, vc_dose_1.0)
t_test_result2 <- t.test(vc_dose_0.5, vc_dose_2.0)
t_test_result3 <- t.test(vc_dose_1.0, vc_dose_2.0)

# Perfome the t-test OJ
t_test_result4 <- t.test(oj_dose_0.5, oj_dose_1.0)
t_test_result5 <- t.test(oj_dose_0.5, oj_dose_2.0)
t_test_result6 <- t.test(oj_dose_1.0, oj_dose_2.0)


#print test result
print(t_test_result1)
print(t_test_result2)
print(t_test_result3)
print(t_test_result4)
print(t_test_result5)
print(t_test_result6)

```

**OJ vs VC**
Based on the calculated p-value of approximately 0.06063 (assuming a significance level of 0.05), we fail to reject the null hypothesis. This indicates that there isn't enough evidence to suggest a significant difference between the lengths of the two groups ("OJ" and "VC"). Therefore, we do not have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths.

**VC 0.5 vs 1.0**
Based on the calculated p-value of approximately 6.811e-07 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (0.5 and 1.0) within the "VC" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

**VC 0.5 vs 2.0**
Based on the calculated p-value of approximately 4.682e-08 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (0.5 and 2.0) within the "VC" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

**VC 1.0 vs 2.0**
Based on the calculated p-value of approximately 9.156e-05 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (2.0 and 1.0) within the "VC" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

**OJ 0.5 vs 1.0**
Based on the calculated p-value of approximately 8.785e-05 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (0.5, 1.0) within the "OJ" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

**OJ 0.5 vs 2.0**
Based on the calculated p-value of approximately 1.324e-06 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (0.5, 2.0) within the "OJ" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

**OJ 1.0 vs 2.0**

Based on the calculated p-value of approximately 0.0392 (which is significantly less than the significance level of 0.05), we reject the null hypothesis. This suggests that there is enough evidence to indicate a significant difference between the lengths of the two dose levels (1.0, 2.0) within the "OJ" group. Therefore, we do have sufficient statistical evidence to conclude that the supplement type significantly affects the lengths at different dose levels.

